---
title: "Data Viz Project"
author: "Stephen Rozen"
date: "2025-03-29"
output:
  html_document: default
  pdf_document: default
knitr:
  opts_chunk: 
    echo: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DescTools)
library(tidyverse)
library(data.table)
library(plotly)
library(broom)
library(brms)
library(tidyverse)



```

```{r}
by_sector <- read.csv("job_postings_by_sector_US.csv")

by_metro <- read.csv("metro_job_postings_us.csv")

cbsa <- read.csv("cbsa-est2023-alldata.csv") %>%
  filter(LSAD == "Metropolitan Statistical Area" | LSAD == "Micropolitan Statistical Area") %>%
  select(POPESTIMATE2023, CBSA, NAME)
  
```

# Will Alex Find a Good Job??

My friend Alex is a third year math major, and goddammit, he can't find a good job for the summer.

This weird because, in addition to being very charming and good-looking, Alex is one of the smartest people I know. Here is a GGplot of Alex, just to give you an idea:

In discussing his desperate joblessness, Alex put forward the idea that the job market *is just really hard right now* for college kids and recent college graduates. I was skeptical, this seems like a classic "grass is greener" complaint. Everybody thinks that the current age is worse than the previous age. Still it is an intriguing question, and when we asked google if the market is hard for college kids, it turned up the fact that \~40% of recent college grads are **under**employed.

This exploration of employment will try to get to the bottom of this dinner-table argument, and also answer some other questions about longer term prospects for college graduates. Are college degrees worthless? Should I start applying to work at CINNABON? How do Alex's particular demographic and educational characteristics affect his chances of finding gainful employment?

## The current conditions -- not too hot

The first place to look is the job postings numbers in fields where Alex is applying for jobs. A recent decline in the number of postings would be a clear indicator that the market is tough. And... It's pretty stark

#### The number of job listings in quantitative fields has been steadily declining since 2022

```{r}


sector_plot <- by_sector %>%
  filter(display_name %in% c("Mathematics", "Scientific Research & Development", "Information Design & Documentation", "Banking & Finance")) %>%
  mutate(date = as.Date(date)) %>%
  ggplot(aes(x = date, y = indeed_job_postings_index, color = display_name)) + 
  geom_line() +
  geom_point(aes(text = paste("Date:", format(date, "%Y-%m-%d"), 
                          "<br>Job Index:", indeed_job_postings_index,
                          "<br>Sector:", display_name)),
             size = 0.1, alpha = 0) + ##error in ggplot aes for some reason
  theme_minimal() + 
  scale_color_OkabeIto() + 
  labs(y = "Job Index",
       x = "Date",
       color = "Sector"
       )+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

ggplotly(sector_plot, tooltip = "text") %>%
  layout(
    legend = list(
      orientation = "h",  
      x = 0.5,            
      xanchor = "center",
      y = 1.1 
    )
  )


```

Yikes. But these are aggregate numbers, maybe there are regional differences that paint a less drastic picture. Alex is a UVA student from NYC, so the two markets where he is primarily conducting his search are New York and Washington DC.

```{r}
metro_plot_data <- by_metro %>%
  rename(CBSA = cbsa_code) %>%
  mutate(date = as.Date(date)) %>%
  filter(date >= as.Date("2025-04-01")) %>%
  group_by(CBSA) %>%
  summarize(index = mean(indeed_job_postings_index)) %>%
  right_join(cbsa) %>%
  arrange(desc(POPESTIMATE2023))
metro_plot_data


        
```

```{r}

cleaned_names <- c(
  "New York, NY",
  "Los Angeles, CA",
  "Chicago, IL",
  "Dallas, TX",
  "Houston, TX",
  "Atlanta, GA",
  "Washington, DC",
  "Philadelphia, PA",
  "Miami, FL",
  "Phoenix, AZ"
)



metro_plot <- metro_plot_data %>%
  mutate(rel_index = index - 100) %>%
  slice(1:10) %>%
  mutate(plt_names = cleaned_names,
         fill_col = ifelse(plt_names %in% c("New York, NY", "Washington, DC"), 
                        "red",
                        "blue")
         )%>%
  ggplot(aes(x = fct_reorder(cleaned_names, rel_index, .desc = TRUE), 
             y = rel_index)) +
  geom_bar(stat="identity", aes(fill = fill_col)) + 
  scale_fill_identity() +

  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        panel.grid.major.x = element_blank())  +
  labs(caption = "indeed.com job postings in april 2025, relative to prepandemic",
       y = "Relative Job Index",
       x = "Metro Area")
  

metro_plot
```

Unfortunately, DC and NYC are currently among the biggest losers among the major metro areas in the US. So, is Alex right about the tough job market? Yes, absolutely (and, really, I shouldn't be surprised, he is usually right).

But I want to delve deeper. Alex may be in a tough spot at the moment, but how are his longer-term prospects? He's a smart dude with a good degree, so it is hard to believe that he will be unable to find gainful employment when we leave UVA. In the next section, we will look at the factors contributing to the employment of recent college graduates.

## Who is gainfully employed... and who isn't?

```{r}
# NOTE: To load data, you must download both the extract's data and the DDI
# and also set the working directory to the folder with these files (or change the path below).

if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

ddi <- read_ipums_ddi("usa_00003.xml")
data <- read_ipums_micro(ddi)
  #select(-SERIAL, -CBSERIAL, -CLUSTER, -STRATA, -SAMPLE, -HHWT, -PERNUM, -PERWT)

samp <- sample_n(data, 1e6, replace = F)
samp_small <- sample_n(data, 1e5, replace = F)

```


```{r}

underemployment <- data %>%
  group_by(OCC2010, YEAR) %>%
  summarise(median(EDUCD), 
            hmed = median(EDUCD) > 100, 
            mod_med = median(EDUCD) > 69,
            mode = max(DescTools::Mode(EDUCD)), 
            hmode = max(DescTools::Mode(EDUCD)) > 100,
            mod_mode = max(DescTools::Mode(EDUCD)) > 69) %>%
  mutate(High_Skill = ifelse(hmed & hmode, T, F),
         Moderate_Skill = ifelse(mod_med & mod_mode, T, F))##Occupation is Modal and Median at least a bachelors/1+ yr college

underemployment ##includes underemployed and unemployed (9920)
data_new <- left_join(data, underemployment)
  data_new

```

```{r}


Unemployment_base <- data %>%
  mutate(young = as.logical(AGE > 21 & AGE < 27), #young or not
         has_bachelors = EDUCD > 100) %>% #has bach or not
  filter(EMPSTAT %in% c(1, 2))  # employed (1) or unemployed (2), excludes NIL/Unknown
  
recent_grads_un <- Unemployment_base %>%
  group_by(YEAR, young, has_bachelors) %>% 
  summarize(recent_grads_rate = 1 - sum(EMPSTAT == 1)/n()) %>%
  filter(young == TRUE & has_bachelors == TRUE) %>%
  select(-has_bachelors, -young)

bach_un <- Unemployment_base %>%
  group_by(YEAR, has_bachelors) %>% 
  summarize(bach_rate = 1 - sum(EMPSTAT == 1)/n()) %>%
  filter(has_bachelors == TRUE) %>%
  select(-has_bachelors)

all_un <- Unemployment_base %>%
  group_by(YEAR) %>% 
  summarize(all_rate = 1 - sum(EMPSTAT == 1)/n())
  


Unemployment_rates <- left_join(recent_grads_un, bach_un) %>%
  left_join(all_un) %>%
  ungroup() %>%
  pivot_longer(c(recent_grads_rate, bach_rate, all_rate)) %>%
  rename(demo_group = name) %>%
  mutate(plot_label = case_when(demo_group == "recent_grads_rate" ~ "Recent College Graduates",
                                demo_group == "bach_rate" ~ "All College Graduates",
                                demo_group == "all_rate" ~ "All Workers",
  )
         )
Unemployment_rates

write.csv(Unemployment_rates, "Unemployment_rates.csv", row.names = F)
```

```{r}
p <- Unemployment_rates %>%
  ggplot(aes(x=YEAR, 
             y = value*100, 
             color = plot_label)) +
  geom_line(size = 1) +
  scale_x_continuous(breaks = seq(min(Unemployment_rates$YEAR), 
                                  max(Unemployment_rates$YEAR), by = 3)) +
  scale_y_continuous(limits = c(0, NA)) +
  theme_minimal() + 
  theme(panel.grid.major.x = element_blank(),  
        panel.grid.minor.x = element_blank(), 
        legend.position = "top") +
  labs(x = "Year",
       y = "Unemployment Rate (%)",
       caption = "source: IPUMS cencus data, all USA") +
  scale_color_manual(name = "",
                     values = c("All Workers" = "#E69F00", 
                               "All College Graduates" = "#56B4E9",
                               "Recent College Graduates" = "#009E73"))


p
```


Unemployment doesn't seem like a big problem, outside of recession times the unemployment rate for recent grads is in range of the broader labor market, around 4%. People are able to find jobs, its the quality of the jobs that is at issue.

Alex, and everyone else entering the real world for the first time wants a GOOD job, not just a job. If all else failed, he could undoubtedly find a job in the service industry (he's very personable!). But the ideally he would put that big, college educated brain to work.

The graph below plots *under*employment, which is the rate of employment in fields that are "below" one's educational attainment. Underemployment is defined here as working in an occupation where the median or modal education level of the employees is less than a bachelors degree. (occupational underemp paper)

#### Almost half of recent college graduates are underemployed

```{r}
data_new_skill_clean <- data_new[!(is.na(data_new$High_Skill) | is.na(data_new$Moderate_Skill)), ] %>%
  mutate(young = as.logical(AGE > 21 & AGE < 27), #young or not
         has_bachelors = EDUCD > 100) %>% #has bach or not
  filter(EMPSTAT %in% c(1, 2))  #employed or unemployed
 ##removes nas in high skill

data_new_skill_clean <- read.csv("data_new_skill_clean.csv")

Underemployment_rates <- data_new_skill_clean %>%
  rename(recent_grad = young) %>%
  group_by(YEAR, recent_grad) %>%
  filter(EDUCD > 100) %>% # at least bachelors
  summarize(strict_underemployment_measure = 1 - sum(High_Skill)/n(), 
            lax_underemployment_measure = 1 - sum(Moderate_Skill)/n()) #

write.csv(Underemployment_rates, "Underemployment_rates.csv", row.names = FALSE)

##DECIDE ABOUT EMPSTAT!!
```

```{r}

p <- Underemployment_rates %>%
  ggplot(aes(x=YEAR, 
             y = strict_underemployment_measure * 100, 
             color = recent_grad)) +
  geom_line(size = 1) +
  scale_x_continuous(breaks = seq(min(Underemployment_rates$YEAR), 
                                  max(Underemployment_rates$YEAR), by = 3)) +
  scale_y_continuous(limits = c(0, NA)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),  
        panel.grid.minor.x = element_blank(),
        legend.position = "top") +
  labs(x = "Year",
       y = "Underemployment Rate (%)",
       caption = "source: IPUMS cencus data, all USA") +
  scale_color_manual(name = "",
                     values = c("FALSE" = "#E69F00", 
                               "TRUE" = "#56B4E9"),
                     labels = c("FALSE" = "All College Graduates", 
                                "TRUE" = "Recent College Graduates"))

p



```

That is a lot of degrees that are not being put to use. Perhaps it's an oversupply issue, though scholarship is mixed, with some analyses finding there may be an under-supply of graduates relative to demand (Goldin & Katz, 2008). Others claim that the quality of higher education is to blame, that "colleges grant baccalaureates to many students who lack basic numeracy and literacy skills and who therefore cannot perform demanding tasks in the workplace".

Whatever the cause, it is notable that underemployment rates are fairly stable over time, again with the caveat of recessions.

As with most of society, race is a factor. Stratifying recent graduate underemployment by ethnic category, we see what one might expect: being part of a marginalized group is a risk factor for being underemployed.

These odds ratios were acquired using a logistic regression, and control for state of residence, degree field, and survey year.

#### Black, Multiracial, and Native college graduates are more likely than Whites to be underemployed even after adjusting for geography and degree field

```{r}
#ODDS RATIO


Underemployment_odds_base <- data_new_skill_clean %>%
  mutate(recent_grad = as.logical(AGE>21 & AGE<27),
         racial_cat = case_when(RACE %in% c(4,5,6) ~ "Asian or Pacific Islander",
                                RACE == 1 ~ "White",
                                RACE == 2 ~ "Black",
                                RACE == 3 ~ "Native",
                                RACE %in% c(8,9) ~ "Multiracial",
                                RACE == 7 ~ "Other")
         )%>%
  filter(EDUCD > 100,              # at least bachelors
         EMPSTAT %in% c(1,2),      #employed or unemployed
         YEAR > 2018) %>%
  mutate(racial_cat = relevel(factor(racial_cat), ref = "White"))

   

# Logistic Regression Model
model <- Underemployment_odds_base %>%
  filter(RACE != 7) %>% ## remove people who didn't report their race
  glm(1-High_Skill ~ racial_cat + STATEICP + DEGFIELD + YEAR, family = binomial, data = .)
model #controls for State, major, survey year


# Extract ORs and CIs
results <- tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term != "(Intercept)")
print(results)

term_fix <- c("Asian or Pacific Islander", "Black", "Multiracial", "Native" )
results <- results %>%
  filter(startsWith(term, "racial_cat")) %>%
  mutate(term = term_fix)

results

```

```{r}

  

plt <- ggplot(results, aes(x = estimate, y = term, 
                    xmin = conf.low, xmax = conf.high)) +
  
  scale_y_discrete(expand = expansion(add = c(0, 1.7))) + 
  
 
  geom_point(color = "blue", size = 2) +
  geom_errorbarh(height = .2, color = "blue") +  
  
 
  scale_x_continuous(
    limits = c(.84, 1.5),
    breaks = c(.85, 1, 1.15, 1.3, 1.45),
    labels = c(".85x as likely", "Just as likely", "1.15x as likely", "1.3x as likely", "1.45x as likely")
  ) +
  
  
  
  theme_minimal() +
  theme(
    axis.line.x = element_blank(),
    panel.grid.major.y = element_blank(), 
    panel.grid.major.x = element_blank(),   
    panel.grid.minor.x = element_blank(),  
    axis.text = element_text(color = "grey30", size = 10),  
    axis.ticks.y = element_line(color = "gray"),  
    axis.title.x = element_text(color = "gray"),  
    axis.title.y = element_blank(),  
    plot.margin = margin(t = 1, r = 20, b = 20, l = 20),
    aspect.ratio = .5,
    plot.title = element_text(size = 20, hjust = .5)
  ) +
  
  labs(
    x = "Likelihood of being underemployed compared to White ethnicity",
  ) +
  

  annotate("text", x = 1, y = length(unique(results$term)) + 1, 
           label = "Comparison group: \nWhite", 
           size = 3.5,
           hjust = .5) +
 


  
  # limit the vertical line height 
  geom_segment(
    x = .85, xend = .85,        
    y = 0.5, yend = length(unique(results$term)) + 0.5,  
    linetype = "solid", color = "gray", linewidth = .2
  ) +
  geom_segment(
    x = 1, xend = 1,        
    y = 0.5, yend = length(unique(results$term)) + 0.5,  
    linetype = "solid", color = "gray", linewidth = 1.2
  ) +
  geom_segment(
    x = 1.15, xend = 1.15,        
    y = 0.5, yend = length(unique(results$term)) + 0.5, 
    linetype = "solid", color = "gray", linewidth = .2
  )  +
  geom_segment(
    x = 1.3, xend = 1.3,        
    y = 0.5, yend = length(unique(results$term)) + 0.5,  
    linetype = "solid", color = "gray", linewidth = .2
  ) +
  geom_segment(
    x = 1.45, xend = 1.45,        
    y = 0.5, yend = length(unique(results$term)) + 0.5, 
    linetype = "solid", color = "gray", linewidth = .2
  ) +
  
  annotate("text", x = 1.15, y = length(unique(results$term)) + 1, 
           label = "Increased risk of \nunderemployment", 
           size = 3.5,
           hjust = 0, 
           color = "grey30") + 

  # Adding an arrow as annotation
  geom_segment(
    x = 1.15, y = length(unique(results$term)) + .55,          
    xend = 1.45,        
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),  
    color = "grey30"
  ) 

plt

```

```{r}
## recent grad underemployment by major


majors <- data_new_skill_clean %>%
  filter(YEAR > 2018,
         has_bachelors == TRUE,
         young = TRUE) %>%
  
  mutate(degree_cat = case_when(
    DEGFIELD %in% c(11, 13) ~ "Natural Resources",
    DEGFIELD %in% c(15, 22, 26, 33, 34, 35, 48, 49, 60) ~ "Arts and Humanities",
    DEGFIELD %in% c(36, 37, 50, 61) ~ "Math and Science",
    DEGFIELD %in% c(21, 24, 25, 51, 57, 58, 59, 38) ~ "Engineering and Tech",
    DEGFIELD %in% c(14, 56) ~ "Architecture",
    DEGFIELD %in% c(29, 52, 54, 55, 64) ~ "Social Sciences",
    DEGFIELD %in% c(32, 53, 62) ~ "Business and Law",
    DEGFIELD %in% c(23, 41, 19, 20, 40) ~ "Iinterdisciplinary")) %>%
  mutate(degree_cat = fct_relevel(degree_cat, "Other", after = Inf))

  
  
  
  
```

```{r}
by_DEGFIELDD <- majors %>%
  group_by(DEGFIELDD, DEGFIELD, degree_cat, STATEICP) %>%
  summarize(
    strict_underemployment_measure = 1 - sum(High_Skill) / n(),
    median_income = median(INCWAGE_CPIU_2010*1.47),
    n = n(),
    .groups = "drop"
  )
by_DEGFIELDD

by_DEGFIELDD_labelled <- as.data.frame(ddi$var_info$val_labels[17]) %>%
  full_join(by_DEGFIELDD, by = join_by(val == DEGFIELDD)) %>%
  rename(DEGFIELDD_label = lbl,
         DEGFIELDD = val) %>%
  full_join(as.data.frame(ddi$var_info$val_labels[16]), 
            by = join_by(DEGFIELD == val))%>%
  rename(DEGFIELD_label = lbl) %>%
  filter(as.numeric(DEGFIELD) > 0 & as.numeric(DEGFIELDD) > 0) %>%
  full_join(as.data.frame(ddi$var_info$val_labels[2]), by = join_by(STATEICP == val)) %>%
  mutate(region = tolower(lbl))
  
write.csv(by_DEGFIELDD_labelled, "by_DEGFIELDD_labelled.csv", row.names = F)


```

It is also evident that people with certain degrees are more likely to be underemployed or underpaid than others. If you're curious about those differences or about your own prospects, use the applet below to explore different degree fields and majors.

```{r}
library(shiny)
library(colorblindr)
library(plotly)
library(tidyverse)
library(maps)
states <- map_data("state")

ui <- fluidPage(
  fluidRow(
    column(12,
           wellPanel(
             fluidRow(
               column(4,
                      
                      selectInput(inputId = "deg_cat",
                                  label = "Degree Category",
                                  choices = c("Natural Resources", 
                                              "Arts and Humanities", 
                                              "Math and Science",
                                              "Engineering and Tech",
                                              "Architecture",
                                              "Social Sciences",
                                              "Business and Law",
                                              "interdisciplinary")
                                  
                                  
                      )
               ),
               
               column(4,
                      
                      selectInput("major", "Pick A Major, Map Their Income", choices = NULL)
               )
               
             )
           )
    )
  ),
  fluidRow(
    column(6,
           plotlyOutput("plot")
    ),
    column(6,
           plotOutput("map"))
  )
)

server <- function(input, output, session) {
  by_DEGFIELDD <- read.csv("by_DEGFIELDD_labelled.csv")
  to_plot <- reactive({
   
      by_DEGFIELDD %>%
        filter(degree_cat == input$deg_cat) #filters to selected category
    
  })
  
  to_dotplot <- reactive({
    by_DEGFIELDD %>%
    filter(degree_cat == input$deg_cat) %>%
    group_by(DEGFIELDD_label) %>%
    summarise(income = mean(median_income),
              underemployment = mean(strict_underemployment_measure))
  })
  
  output$plot <- renderPlotly({
    
    
    
    # Then create the plot with the summarized data
    p <- to_dotplot() %>%
      ggplot(aes(x = income, y = underemployment,
                       text = paste(
                         "Major: ", DEGFIELDD_label,  
                         "<br>Underemployment Rate: ", round(underemployment, digits = 3),
                         "<br>Median Income: $", round(income, 0)
                       ))) +
      geom_point(color = "#009E73") +
      labs(x = "Median Income ($)",
           title = paste0("Income and Employment for Majors in\n", input$deg_cat)) +
      scale_x_continuous(limits = c(20000, 115000),
                         breaks = c(20000, 40000, 60000, 80000, 100000),
                         labels = c("20k", "40k", "60k", "80k", "100k")) +
      scale_y_continuous(
        name = "Underemployment Rate"
      ) +
      theme_minimal() + 
      theme(plot.title = element_text(size = 12, hjust = .5))
    
    ggplotly(p, tooltip = "text")

    
  })
  
  
  observe({
    
    major_choices <- to_plot() %>%
      select(DEGFIELDD_label) %>%
      distinct() %>%
      pull()
    
    updateSelectInput(session, "major",
                      choices = major_choices,
                      selected = NULL)
  })
  
  
  output$map <- renderPlot({
    req(input$major)
    choro_data <- to_plot() %>%
      filter(DEGFIELDD_label == input$major)
    
    choro <- left_join(choro_data, states, by = join_by(region == region))
    print(choro)
    
    ggplot(choro, aes(long, lat, group = group, 
                      fill = median_income)) +
      geom_polygon() +
      coord_quickmap() +
      labs(fill = "Median\nIncome ($)",
           title = paste0(input$major, " Majors' Median Income")) +
      theme_minimal() +
      theme(legend.position = "inside",
            legend.position.inside = c(.90, .2),
            plot.title = element_text(size = 15, hjust = .5),
            axis.line.x = element_blank(),
            panel.grid.major.y = element_blank(), 
            panel.grid.major.x = element_blank(),   
            panel.grid.minor.x = element_blank(),
            panel.grid.minor.y = element_blank(),
            axis.text = element_blank(),
            axis.title = element_blank())
    
    
  })
  
}
shinyApp(ui = ui, server = server)
```






Let us return to Alex though, and see how people like him are doing. He has a degree in Math and a double major in History. He is white and male, and a native English speaker. Alex's traits seem like they would put him at an advantage in the job market. The following animation looks at how people with similar characteristics stack up in terms of employment and income. This may give us more insight into how Alex himself will fare.

Here is a breakdown for this animation:

\- **engl_white**: Now here are the [English-speaking white folks]{style="color:red;"}, the majority group. They're doing pretty well

\- **quants**: Here are our stereotypical quants, [the dudes studying math]{style="color:red;"} -- not too shabby nerds!!

\- **alex_like**: These are the people most [similar to Alex]{style="color:red;"} -- recent graduates with double majors and no Masters or PHD. If Alex is anything like the folks, it seems like he'll probably find a decent job.

```{r}
## People in the dataset of last 5 years matching kurtz's description

income_modelling_base <- Underemployment_odds_base %>%
  mutate(double_major = as.logical(DEGFIELD2 != 0), ##add a T/F for whether person is double major
         graduate_degree = as.logical(EDUCD > 113),
         DEGFIELD = as.factor(DEGFIELD),
         LANGUAGE = as.logical(LANGUAGE == 1)
         ) %>%
  filter(EMPSTAT == 1)
income_modelling_base
```

```{r}




alex_subsets <- income_modelling_base %>%
  group_by(SEX, LANGUAGE, DEGFIELD, recent_grad, racial_cat, double_major, graduate_degree) %>% #add hispan?
  summarize(
    strict_underemployment_measure = 1 - sum(High_Skill) / n(),
    med_income = median(INCWAGE_CPIU_2010) * 1.47,
    n = n()
    ) %>%
  mutate(engl_white = ifelse(LANGUAGE == 1 & racial_cat == "White", TRUE, FALSE)) %>%
  mutate(quants = ifelse(engl_white == 1 & SEX == 1 & DEGFIELD == 37, TRUE, FALSE)) %>%
  mutate(alex_like = ifelse(quants == 1 & recent_grad == TRUE &
         double_major == TRUE &
         graduate_degree == FALSE, TRUE, FALSE)
         ) %>%
  left_join(as.data.frame(ddi$var_info$val_labels[16]), 
            by = join_by(DEGFIELD == val))
 

alex_subsets <- alex_subsets %>%
 cbind(seq(1, nrow(alex_subsets))) %>%
 rename(id = ...15) %>%
 pivot_longer(
    cols = c(engl_white, quants, alex_like),
    names_to = "frame",
    values_to = "is_red",
    values_drop_na = TRUE)
     
write.csv(alex_subsets, "alex_subsets.csv", row.names = F)


```

```{r}

p1 <- alex_subsets %>%
  filter(n>5) %>%
  ggplot(aes(x = med_income, 
             y = strict_underemployment_measure, 
             color = is_red,
             text = paste(
            "<br>Underemployment Ratio: ", round(strict_underemployment_measure, digits = 3),
            "<br>Median Income: $", med_income, ##account for change in currency value
            "<br>Degree Field:", lbl,  
            "<br>Masters/PHD: ", graduate_degree,
            "<br>Recent College Graduate: ", recent_grad,
            "<br>Ethnicity: ", racial_cat,
            "<br>Sex: ", ifelse(SEX == 1, "Male", "Female"),
            "<br>N: ", n
            
          )
          )) +
  geom_jitter(aes(frame = fct_relevel(frame, "engl_white", "quants", "alex_like"))) +
  scale_x_continuous(name = "Median Income ($)",
                     limits = c(0, 180000),
                     breaks = c(0, 30000, 60000, 90000, 120000, 150000, 180000),
                     labels = c("0", "30k", "60k", "90k", "120k", "150k", "180k")) +
  scale_y_continuous(
    name = "Underemployment Ratio"  # Set the y-axis title
  ) +
  theme_minimal() +
  scale_color_manual(
    values = c("grey", "red")) +
  theme(legend.position = "none")

ggplotly(p1, tooltip = "text") %>%
  animation_opts(
    1000, easing = "elastic", redraw = FALSE
  ) %>%
  animation_slider(
    currentvalue = NULL
  )
```



Finally, lets look at income distribution. We all want to make money, and factors that we have explored so far like demographics and educational attainment are somewhat predictive of our future earnings. Emphasis on the *somewhat*, the model I fitted with millions of data points had an R\^2 = .11. We can't all be perfect.

The applet below will draw a prediction interval for your future income based on the characteristics you input. Think of it as a very very fuzzy crystal ball🔮
```{r}
## THIS IS A REPEAT OF A BUNCH OF CODE
## using a random sample of the dataset, in order to make a smaller model


sample <- sample_n(data, 100000)

underemployment <- samp %>%
  group_by(OCC2010, YEAR) %>%
  summarise(median(EDUCD), 
            hmed = median(EDUCD) > 100, 
            mod_med = median(EDUCD) > 69,
            mode = max(DescTools::Mode(EDUCD)), 
            hmode = max(DescTools::Mode(EDUCD)) > 100,
            mod_mode = max(DescTools::Mode(EDUCD)) > 69) %>%
  mutate(High_Skill = ifelse(hmed & hmode, T, F),
         Moderate_Skill = ifelse(mod_med & mod_mode, T, F))##Occupation is Modal and Median at least a bachelors/1+ yr college

underemployment ##includes underemployed and unemployed (9920)
data_new <- left_join(samp, underemployment)


data_new_skill_clean <- data_new[!(is.na(data_new$High_Skill) | is.na(data_new$Moderate_Skill)), ] %>%
  mutate(young = as.logical(AGE > 21 & AGE < 27), #young or not
         has_bachelors = EDUCD > 100) %>% #has bach or not
  filter(EMPSTAT %in% c(1, 2)) 

Underemployment_odds_base <- data_new_skill_clean %>%
  mutate(recent_grad = as.logical(AGE>21 & AGE<27),
         racial_cat = case_when(RACE %in% c(4,5,6) ~ "Asian or Pacific Islander",
                                RACE == 1 ~ "White",
                                RACE == 2 ~ "Black",
                                RACE == 3 ~ "Native",
                                RACE %in% c(8,9) ~ "Multiracial",
                                RACE == 7 ~ "Other")
         )%>%
  filter(EDUCD > 100,              # at least bachelors
         EMPSTAT %in% c(1,2),      #employed or unemployed
         YEAR > 2018) %>%
  mutate(racial_cat = relevel(factor(racial_cat), ref = "White"))

income_modelling_base <- Underemployment_odds_base %>%
  mutate(double_major = as.logical(DEGFIELD2 != 0), ##add a T/F for whether person is double major
         graduate_degree = as.logical(EDUCD > 113),
         DEGFIELD = as.factor(DEGFIELD),
         LANGUAGE = as.logical(LANGUAGE == 1),
         wage_2025 = INCWAGE_CPIU_2010*1.47,
         log_wage_2025 = log(wage_2025)) %>%
  filter(EMPSTAT == 1,                    ##only the employed people
         INCWAGE_CPIU_2010 != 0) %>%          #excludes those with income of 0 
        mutate(age_centered = AGE - mean(AGE)) %>%
  mutate(SEX = factor(SEX), 
         LANGUAGE = factor(LANGUAGE), 
         DEGFIELD = factor(DEGFIELD), 
         racial_cat = factor(racial_cat), 
           graduate_degree = factor(graduate_degree))





income_modelling_base


```


```{r}

bayes_model <- brm(
  wage_2025 ~ SEX + LANGUAGE + DEGFIELD + age_centered + racial_cat + graduate_degree,
  data = income_modelling_base,
  family = lognormal(),
  prior = c(
    prior(normal(log(55000), 0.2), class = "Intercept"),       
    prior(cauchy(0, 5), class = "sigma"),                  
    prior(normal(0, 0.5), class = "b") 
  ),
  chains = 3,
  iter = 1000,
  warmup = 500,
  cores = 4,
  seed = 123
)

summary(bayes_model)
plot(bayes_model, variable = c("b_age_centered"))
saveRDS(bayes_model, "bayesreal.RDS")
```

wage_2025 ~ LN(sig, mu)
sig ~ HC(0, 5)
mu ~ normal(log(55000), s)
b ~ normal(0, .5) #median wage multiplier in [exp(-1), exp(1)]



```{r}
bayes_model <- readRDS("bayesfake.RDS")
kurtz_pred <- data.frame(SEX = 1, LANGUAGE = TRUE , DEGFIELD = as.factor(37), age_centered = -19.79, racial_cat = "White", graduate_degree = FALSE)




predict(bayes_model, newdata = kurtz_pred, probs = c(0.10, 0.9))


# 1. Draw predictive samples (posterior predictive = includes model error)
pred_samples <- posterior_predict(bayes_model, newdata = kurtz_pred)

# 2. Flatten to a vector (only 1 row in kurtz_pred)
wage_preds <- pred_samples[,1]

# 3. Compute 90% predictive interval
ci_90 <- quantile(wage_preds, probs = c(0.2, 0.8))

# 4. Plot with ggplot2
library(ggplot2)

ggplot(data.frame(wage = wage_preds), aes(x = wage)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = ci_90, linetype = "dashed", color = "red") +
  annotate("text", x = ci_90[1], y = 0, label = paste0("15%: $", round(ci_90[1])), hjust = -0.1, vjust = -1.5, color = "red") +
  annotate("text", x = ci_90[2], y = 0, label = paste0("85%: $", round(ci_90[2])), hjust = 1.1, vjust = -1.5, color = "red") +
  scale_x_continuous(limits = c(0, 4e5)) + 
  labs(title = "Posterior Predictive Distribution of Wage",
       subtitle = "With 70% Predictive Interval",
       x = "Predicted Wage ($)", y = "Density") +
  geom_point(aes(x = median(wage_preds), y = 0), color = "black", size = 3) +
annotate("text", x = median(wage_preds), y = 0, label = paste0("Median: $", round(median(wage_preds))), vjust = -1) +
  theme_minimal()
```

```{r}
library(tidyverse)
library(shiny)
library(brms)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(width = 5,
                 fluidPage(
                   fluidRow(
                     column(6,
                            selectizeInput(
                              'deg_field_input', 
                              label = "Degree Field", 
                              choices = NULL,
                              options = list(placeholder = 'search degree field'),
                              selected = NULL
                            )
                     ),
                     column(6,
                            actionButton(inputId = "button",
                                         label = "Generate data")
                     )
                   ),
                   fluidRow(
                     column(6,
                            radioButtons(inputId = "sex",
                                         label = "Sex",
                                         choices = c("Male", "Female"))
                     ),
                     column(6, 
                            radioButtons(inputId = "recent_grad",
                                         label = "Recent Graduate",
                                         choices = c("True", "False"))
                     )
                   ),
                   fluidRow(
                     column(6, 
                            radioButtons(inputId = "lang",
                                         label = "English First Language",
                                         choices = c("True", "False"))
                     ),
                     column(6,
                            radioButtons(inputId = "graduate_degree",
                                         label = "Has Graduate Degree",
                                         choices = c("True", "False"))
                     )
                   ),
                   fluidRow(
                     column(6, 
                            selectInput(inputId = "race",
                                        label = "Racial Category",
                                        choices = c("Black", "Multiracial", "Native", 
                                                    "White", "Asian or Pacific Islander",
                                                    "Other"))
                     ),
                     column(6, 
                            numericInput("age",
                                         "Age", 
                                         value = 40.7))
                   )
                 )
    ),
    mainPanel(width = 7,
      plotOutput("plot")

    )
  )
)

server <- function(input, output, session) {
  Labelled_DEGFIELD <- read.csv("Labelled DEGFIELD.csv")
  choices <- Labelled_DEGFIELD$lbl
  income_model <- readRDS("bayesreal.RDS")
  
  
  observe({
    updateSelectizeInput(session, 'deg_field_input', 
                         choices = choices, 
                         server = TRUE,
                         selected = character(0))
  })
  
  user_params <- eventReactive(input$button, {
    df <- data.frame(SEX = ifelse(input$sex == "Male", 1, 2),
                    LANGUAGE = ifelse(input$lang == "True", T, F),
                    recent_grad = ifelse(input$recent_grad == "True", T, F),
                    graduate_degree = ifelse(input$graduate_degree == "True", T, F), 
                    racial_cat = input$race,
                    age_centered = input$age - 40.798,
                    DEGFIELD = Labelled_DEGFIELD %>% 
                      filter(lbl == input$deg_field_input) %>%
                      pull(val)  
                    
    )
    
    
  })
  
  
  
  predictions_data <- reactive({
  req(input$button)
  #predictions <- predict(income_model, newdata = user_params(), interval = "prediction", level = 0.80)
  pred_samples <- posterior_predict(income_model, newdata = user_params())
  wage_preds <- pred_samples[,1]
  ci_90 <- quantile(wage_preds, probs = c(0.2, 0.8))
  
  list(wage_preds = wage_preds,
       ci_90 = ci_90)
  })
  
    
  output$plot <- renderPlot({
    req(predictions_data())
    
    pdata <- predictions_data()
    ci_90 <- pdata$ci_90
    wage_preds <- pdata$wage_preds
    
    dens <- density(wage_preds)

# Convert to data frame for plotting and exact matching
  dens_df <- data.frame(x = dens$x, y = dens$y)

# Find the density y-value closest to the median x-value
  q_x <- quantile(wage_preds, .001)
  q_y <- approx(dens$x, dens$y, xout = q_x)$y
    
    

    
    ggplot(data.frame(wage = wage_preds), aes(x = wage)) +
    geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = ci_90, linetype = "dashed", color = "red") +
  annotate("text", x = ci_90[1], y = 0, label = paste0("15% \n$", round(ci_90[1])), hjust = 1.1, vjust = -1.8, color = "red") +
  annotate("text", x = ci_90[2], y = 0, label = paste0("85% \n$", round(ci_90[2])), hjust = -0.1, vjust = -1.8, color = "red") +
  labs(title = "Posterior Predictive Distribution of Wage",
       subtitle = "With 70% Predictive Interval",
       x = "Predicted Wage ($)", y = "Probability Density") +
  scale_x_continuous(
                         limits = c(0, 310000),
                         breaks = c(seq(0, 3e5, 3e4)),
                         labels = c("0", "30k", "60k", "90k", "120k", "150k", "180k", "210k", "240k", "270k", "300k"))  +
  geom_point(aes(x = median(wage_preds), y = 0), color = "black", size = 3) +
annotate("text", x = median(wage_preds), y = 0, label = paste0("Median: $", round(median(wage_preds))), vjust = -1) +
  theme_minimal() +
      theme(axis.text.y = element_blank()) 
    
  
  
  }
  )
  

  
  
  
}

shinyApp(ui = ui, server = server)
```
## So will Alex find a job?

The prospects for Alex in the medium-term, that is beyond just this year, are good. People with his education and background have fairly low probability of being underemployed. Though the job market is a bit shaky at the moment, it shows long term consistency, so he should be confident that it will return to normal. What's more, underemployment appears to be trending downward, and knowledge work is expected to grow over the next decade (Nguyen). And finally, on a personal level, I know that employers will find his wealth of smarts and creativity very attractive.

**References**:

Attewell, Paul, and Dirk Witteveen. "Occupational Marginalization, Underemployment, and Earnings Inequality among College Graduates." *Research in Social Stratification and Mobility*, vol. 83, Feb. 2023, p. 100761. *ScienceDirect*, <https://doi.org/10.1016/j.rssm.2023.100761>.

Nguyen, An, et al. "New BLS Employment Projections: 3 Charts." *DOL Blog*, <http://blog.dol.gov/2024/09/06/new-bls-employment-projections-3-charts>. Accessed 19 Apr. 2025.

Data from:

Ruggles, Steven, et al. *IPUMS USA: Version 16.0*. 16.0, Minneapolis, MN: IPUMS, 2025. *DOI.org (Datacite)*, <https://doi.org/10.18128/D010.V16.0>.

*Hiring-Lab/Job_postings_tracker*. 2020. Hiring Lab, 16 Apr. 2025. *GitHub*, <https://github.com/hiring-lab/job_postings_tracker>.



